A Logical Solution of the Linearity Challenge

### 2025-11-30 paper/Context.lagda.tex
new notation for splitting

\begin{code}[hide]
infix 4 _≃_+_ _≃_,_
\end{code}
\begin{code}
data _≃_+_ : Context → Context → Context → Set where
  split-e  : [] ≃ [] + []
  split-l  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ A ∷ Δ + Θ
  split-r  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ Δ + A ∷ Θ
\end{code}

As an example, below is the splitting of the context $[A, B, C, D]$ into $[A,
D]$ and $[B, C]$. Note how the sequence of \AgdaInductiveConstructor{split-l}
and \AgdaInductiveConstructor{split-r} applications determines where each of the
types in $[A, B, C, D]$ ends up in $[A, D]$ and $[B, C]$.

\begin{code}[hide]
module _ where
  postulate A B C D : Type
\end{code}
\begin{code}
  splitting-example : (A ∷ B ∷ C ∷ D ∷ []) ≃ (A ∷ D ∷ []) + (B ∷ C ∷ [])
  splitting-example = split-l (split-r (split-r (split-l split-e)))
\end{code}

@@

\begin{code}
_≃_,_ : Context → Type → Context → Set
Γ ≃ A , Δ = Γ ≃ [ A ] + Δ
\end{code}

\begin{code}
+-comm    : ∀{Γ Δ Θ} → Γ ≃ Δ + Θ → Γ ≃ Θ + Δ
+-unit-l  : ∀{Γ} → Γ ≃ [] + Γ
+-unit-r  : ∀{Γ} → Γ ≃ Γ + []
\end{code}
\begin{code}[hide]
+-comm split-e = split-e
+-comm (split-l p) = split-r (+-comm p)
+-comm (split-r p) = split-l (+-comm p)

+-unit-l {[]} = split-e
+-unit-l {_ ∷ _} = split-r +-unit-l

+-unit-r = +-comm +-unit-l

++≃+ : ∀{Γ Δ} → Γ ++ Δ ≃ Γ + Δ
++≃+ {[]} = +-unit-l
++≃+ {_ ∷ _} = split-l ++≃+
\end{code}

\begin{code}
+-assoc-r  : ∀{Γ Δ Θ Δ′ Θ′} → Γ ≃ Δ + Θ → Θ ≃ Δ′ + Θ′ →
             ∃[ Γ′ ] Γ′ ≃ Δ + Δ′ × Γ ≃ Γ′ + Θ′
+-assoc-l  : ∀{Γ Δ Θ Δ′ Θ′} → Γ ≃ Δ + Θ → Δ ≃ Δ′ + Θ′ →
             ∃[ Γ′ ] Γ′ ≃ Θ′ + Θ × Γ ≃ Δ′ + Γ′
\end{code}
\begin{code}[hide]
+-assoc-r split-e split-e = [] , split-e , split-e
+-assoc-r (split-l p) q with +-assoc-r p q
... | _ , p′ , q′ = _ , split-l p′ , split-l q′
+-assoc-r (split-r p) (split-l q) with +-assoc-r p q
... | _ , p′ , q′ = _ , split-r p′ , split-l q′
+-assoc-r (split-r p) (split-r q) with +-assoc-r p q
... | _ , p′ , q′ = _ , p′ , split-r q′

+-assoc-l p q with +-assoc-r (+-comm p) (+-comm q)
... | Δ , r , p′ = Δ , +-comm r , +-comm p′

+-empty-l : ∀{Γ Δ} → Γ ≃ [] + Δ → Γ ≡ Δ
+-empty-l split-e = refl
+-empty-l (split-r p) = cong (_ ∷_) (+-empty-l p)

+-sing-l : ∀{A B Γ} → [ A ] ≃ B , Γ → A ≡ B × Γ ≡ []
+-sing-l (split-l split-e) = refl , refl
\end{code}

>>>>

\begin{code}[hide]
infix 4 _≃_+_ _≃_,_
\end{code}
\begin{code}
data _≃_+_ : Context → Context → Context → Set where
  •   : [] ≃ [] + []
  <_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ A ∷ Δ + Θ
  >_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ Δ + A ∷ Θ
\end{code}

When splitting a context $\Context$ into $\ContextD + \ContextE$,
for each type in $\Context$ we use one of the prefix operators \AgdaInductiveConstructor{<\_} and \AgdaInductiveConstructor{>\_} to indicate whether it is meant to be placed in $\ContextD$ or $\ContextE$,
respectively.
%
As an example, below is the splitting of the context $[A, B, C, D]$ into $[B] + [A, C, D]$.

\begin{code}[hide]
module _ where
  postulate A B C D : Type
\end{code}
\begin{code}
  splitting-example₁ : (A ∷ B ∷ C ∷ D ∷ []) ≃ (B ∷ []) + (A ∷ C ∷ D ∷ [])
  splitting-example₁ = > < > > •
\end{code}

\begin{code}
+-comm    : ∀{Γ Δ Θ} → Γ ≃ Δ + Θ → Γ ≃ Θ + Δ
≫         : ∀{Γ} → Γ ≃ [] + Γ
≪         : ∀{Γ} → Γ ≃ Γ + []
\end{code}
\begin{code}[hide]
+-comm • = •
+-comm (< p) = > (+-comm p)
+-comm (> p) = < (+-comm p)

++≃+ : ∀{Γ Δ} → Γ ++ Δ ≃ Γ + Δ
++≃+ {[]}    {[]}    = •
++≃+ {[]}    {_ ∷ _} = > ++≃+
++≃+ {_ ∷ _} {_}     = < ++≃+

≫ = ++≃+ {[]}

≪ = +-comm ≫
\end{code}

\begin{code}
+-assoc-r  : ∀{Γ Δ Θ Δ′ Θ′} → Γ ≃ Δ + Θ → Θ ≃ Δ′ + Θ′ →
             ∃[ Γ′ ] Γ′ ≃ Δ + Δ′ × Γ ≃ Γ′ + Θ′
+-assoc-l  : ∀{Γ Δ Θ Δ′ Θ′} → Γ ≃ Δ + Θ → Δ ≃ Δ′ + Θ′ →
             ∃[ Γ′ ] Γ′ ≃ Θ′ + Θ × Γ ≃ Δ′ + Γ′
\end{code}
\begin{code}[hide]
+-assoc-r • • = [] , • , •
+-assoc-r (< p) q with +-assoc-r p q
... | _ , p′ , q′ = _ , < p′ , < q′
+-assoc-r (> p) (< q) with +-assoc-r p q
... | _ , p′ , q′ = _ , > p′ , < q′
+-assoc-r (> p) (> q) with +-assoc-r p q
... | _ , p′ , q′ = _ , p′ , > q′

+-assoc-l p q with +-assoc-r (+-comm p) (+-comm q)
... | Δ , r , p′ = Δ , +-comm r , +-comm p′
\end{code}

@@

\begin{code}
_≃_,_ : Context → Type → Context → Set
Γ ≃ A , Δ = Γ ≃ [ A ] + Δ
\end{code}
\begin{code}[hide]
+-empty-l : ∀{Γ Δ} → Γ ≃ [] + Δ → Γ ≡ Δ
+-empty-l • = refl
+-empty-l (> p) = cong (_ ∷_) (+-empty-l p)

+-sing-l : ∀{A B Γ} → [ A ] ≃ B , Γ → A ≡ B × Γ ≡ []
+-sing-l (< •) = refl , refl
\end{code}

Note that, when specifying a splitting $\SimpleSplitting\ContextC{A}\ContextE$, it may be convenient
to use \AgdaFunction{≫} as a shortcut for a sequence of applications of \AgdaInductiveConstructor{>\_}
after the $A$ type has been reached. For instance, the same sample splitting shown earlier can be
written equivalently and in a more compact way as shown below. More examples will be given in \Cref{sec:examples}.

\begin{code}[hide]
module _ where
\end{code}
\begin{code}
  splitting-example₂ : (A ∷ B ∷ C ∷ D ∷ []) ≃ (B ∷ []) + (A ∷ C ∷ D ∷ [])
  splitting-example₂ = > < ≫
\end{code}

### 2025-11-30 paper/Permutations.lagda.tex
new notation for splitting

Just like permutations from Agda's standard library, also \AgdaDatatype{\_\#\_}
enjoys a number of expected (and useful) properties, such as symmetry. Among all
properties, in the following we will make use of a specific one that connects
context splitting and permutations: if $\Context \simeq A, \ContextD$, then
$\Context$ must be a permutation of
$A~\AgdaInductiveConstructor{∷}~\ContextD$.\Luca{Ci sono molte altre proprietà
non descritte, controllare quelle che servono davvero.}

\begin{code}
#cons : ∀{A Γ Δ} → Γ ≃ A , Δ → (A ∷ Δ) # Γ
#cons (split-l p) with +-empty-l p
... | refl = #refl
#cons (split-r p) = #tran #here (#next (#cons p))
\end{code}

\begin{code}[hide]
#split : ∀{Γ Γ₁ Γ₂ Δ} → Γ # Δ → Γ ≃ Γ₁ + Γ₂ → ∃[ Δ₁ ] ∃[ Δ₂ ] (Δ ≃ Δ₁ + Δ₂ × Γ₁ # Δ₁ × Γ₂ # Δ₂)
#split #refl p = _ , _ , p , #refl , #refl
#split (#next π) (split-l p) with #split π p
... | Δ₁ , Δ₂ , q , π₁ , π₂ = _ ∷ Δ₁ , Δ₂ , split-l q , #next π₁ , π₂
#split (#next π) (split-r p) with #split π p
... | Δ₁ , Δ₂ , q , π₁ , π₂ = Δ₁ , _ ∷ Δ₂ , split-r q , π₁ , #next π₂
#split #here (split-l (split-l p)) = _ , _ , split-l (split-l p) , #here , #refl
#split #here (split-l (split-r p)) = _ , _ , split-r (split-l p) , #refl , #refl
#split #here (split-r (split-l p)) = _ , _ , split-l (split-r p) , #refl , #refl
#split #here (split-r (split-r p)) = _ , _ , split-r (split-r p) , #refl , #here
#split (#tran π π′) p with #split π p
... | Θ₁ , Θ₂ , p′ , π₁ , π₂ with #split π′ p′
... | Δ₁ , Δ₂ , q , π₁′ , π₂′ = Δ₁ , Δ₂ , q , #tran π₁ π₁′ , #tran π₂ π₂′

@@

#shift : ∀{A Γ Δ} → (Γ ++ A ∷ Δ) # (A ∷ Γ ++ Δ)
#shift {_} {[]} = #refl
#shift {_} {B ∷ Γ} = #tran (#next #shift) #here

+++# : ∀{Γ Γ₁ Γ₂} → Γ ≃ Γ₁ + Γ₂ → (Γ₁ ++ Γ₂) # Γ
+++# split-e = #refl
+++# (split-l p) = #next (+++# p)
+++# (split-r p) = #tran #shift (#next (+++# p))

@@

#un+ : ∀{Γ Γ₁ Γ₂} → Γ ≃ Γ₁ + Γ₂ → Un Γ₁ → Un Γ₂ → Un Γ
#un+ split-e un-[] un-[] = un-[]
#un+ (split-l p) (un-∷ un₁) un₂ = un-∷ (#un+ p un₁ un₂)
#un+ (split-r p) un₁ (un-∷ un₂) = un-∷ (#un+ p un₁ un₂)
\end{code}

>>>>

Just like permutations from Agda's standard library, also \AgdaDatatype{\_\#\_}
enjoys a number of expected (and useful) properties, such as symmetry. Among all
properties, in the following we will make use of a specific one that connects
context splitting and permutations: if $\Splitting\ContextC\ContextD\ContextE$,
then $\ContextC$ is a permutation of the concatenation of $\ContextD$ and
$\ContextE$.\Luca{Ci sono molte altre proprietà non descritte, controllare
quelle che servono davvero.}

\begin{code}[hide]
#split : ∀{Γ Γ₁ Γ₂ Δ} → Γ # Δ → Γ ≃ Γ₁ + Γ₂ → ∃[ Δ₁ ] ∃[ Δ₂ ] (Δ ≃ Δ₁ + Δ₂ × Γ₁ # Δ₁ × Γ₂ # Δ₂)
#split #refl p = _ , _ , p , #refl , #refl
#split (#next π) (< p) with #split π p
... | Δ₁ , Δ₂ , q , π₁ , π₂ = _ ∷ Δ₁ , Δ₂ , < q , #next π₁ , π₂
#split (#next π) (> p) with #split π p
... | Δ₁ , Δ₂ , q , π₁ , π₂ = Δ₁ , _ ∷ Δ₂ , > q , π₁ , #next π₂
#split #here (< < p) = _ , _ , < < p , #here , #refl
#split #here (< > p) = _ , _ , > < p , #refl , #refl
#split #here (> < p) = _ , _ , < > p , #refl , #refl
#split #here (> > p) = _ , _ , > > p , #refl , #here
#split (#tran π π′) p with #split π p
... | Θ₁ , Θ₂ , p′ , π₁ , π₂ with #split π′ p′
... | Δ₁ , Δ₂ , q , π₁′ , π₂′ = Δ₁ , Δ₂ , q , #tran π₁ π₁′ , #tran π₂ π₂′

@@

#shift : ∀{A Γ Δ} → (Γ ++ A ∷ Δ) # (A ∷ Γ ++ Δ)
#shift {_} {[]} = #refl
#shift {_} {B ∷ Γ} = #tran (#next #shift) #here
\end{code}

\begin{code}
#append : ∀{Γ Γ₁ Γ₂} → Γ ≃ Γ₁ + Γ₂ → (Γ₁ ++ Γ₂) # Γ
\end{code}
\begin{code}[hide]
#append • = #refl
#append (< p) = #next (#append p)
#append (> p) = #tran #shift (#next (#append p))

@@

#un+ : ∀{Γ Γ₁ Γ₂} → Γ ≃ Γ₁ + Γ₂ → Un Γ₁ → Un Γ₂ → Un Γ
#un+ • un-[] un-[] = un-[]
#un+ (< p) (un-∷ un₁) un₂ = un-∷ (#un+ p un₁ un₂)
#un+ (> p) un₁ (un-∷ un₂) = un-∷ (#un+ p un₁ un₂)
\end{code}

### 2025-11-30 paper/Process.lagda.tex
added prose for permutations
new notation for splitting

\begin{code}[hide]
open import Data.Bool using (Bool; true; false; if_then_else_)
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong; cong₂)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)

@@

We adopt an \emph{intrinsically typed} representation of processes, so that a
process is also a proof derivation showing that the process is well typed. This
choice increases the effort in the definition of the datatypes for processes and
their operational semantics, but it pays off in the rest of the formalization
for a number of reasons:

@@

\begin{AgdaAlign}
\begin{code}
data Process : Context → Set where
   link      : ∀{Γ A B} (d : Dual A B) (p : Γ ≃ [ A ] + [ B ]) → Process Γ
\end{code}

@@

\begin{code}
   fail      : ∀{Γ Δ} (p : Γ ≃ ⊤ , Δ) → Process Γ
\end{code}

@@

\begin{code}
   close     : Process [ 𝟙 ]
\end{code}

@@

\begin{code}
   wait      : ∀{Γ Δ} (p : Γ ≃ ⊥ , Δ) → Process Δ → Process Γ
\end{code}

@@

\begin{code}
   select    : ∀{Γ Δ A B} (x : Bool) (p : Γ ≃ A ⊕ B , Δ) →
               Process ((if x then A else B) ∷ Δ) → Process Γ
\end{code}

We represent the tags $\InTag_1$ and $\InTag_2$ with Agda′s boolean values
\AgdaInductiveConstructor{true} and \AgdaInductiveConstructor{false}. In this
way we can use a single constructor \AgdaInductiveConstructor{select} for the
selections $\Select\x{\InTag_i}.P$. According to the semantics of \Calculus, the
channel on which the selection is performed is consumed and the continuation
process uses a fresh continuation channel. As a consequence, the type of the
continuation channel, which is either $A$ or $B$ depending on the value of the
tag $x$, is \emph{prepended} in front of the typing context used to type the
continuation process.

\begin{code}
   case      : ∀{Γ Δ A B} (p : Γ ≃ A & B , Δ) →
               Process (A ∷ Δ) → Process (B ∷ Δ) → Process Γ
\end{code}

@@

\begin{code}
   fork      : ∀{Γ Δ Γ₁ Γ₂ A B} (p : Γ ≃ A ⊗ B , Δ) (q : Δ ≃ Γ₁ + Γ₂) →
               Process (A ∷ Γ₁) → Process (B ∷ Γ₂) → Process Γ
\end{code}

@@

\begin{code}
   join      : ∀{Γ Δ A B} (p : Γ ≃ A ⅋ B , Δ) →
               Process (B ∷ A ∷ Δ) → Process Γ
\end{code}

@@

\begin{code}
   server    : ∀{Γ Δ A} (p : Γ ≃ ¡ A , Δ) (un : Un Δ) →
               Process (A ∷ Δ) → Process Γ
\end{code}

@@

\begin{code}
   client    : ∀{Γ Δ A} (p : Γ ≃ ¿ A , Δ) → Process (A ∷ Δ) → Process Γ
   weaken    : ∀{Γ Δ A} (p : Γ ≃ ¿ A , Δ) → Process Δ → Process Γ
   contract  : ∀{Γ Δ A} (p : Γ ≃ ¿ A , Δ) → Process (¿ A ∷ ¿ A ∷ Δ) → Process Γ
\end{code}

@@

\begin{code}
   cut       : ∀{Γ Γ₁ Γ₂ A B} (d : Dual A B) (p : Γ ≃ Γ₁ + Γ₂) →
               Process (A ∷ Γ₁) → Process (B ∷ Γ₂) → Process Γ
\end{code}

@@

\end{AgdaAlign}

\begin{code}
#process : ∀{Γ Δ} → Γ # Δ → Process Γ → Process Δ
\end{code}
\begin{code}[hide]
#process π (link d p) with #one+ π p
... | Δ′ , q , π′ with #singleton-inv π′
... | refl = link d q
#process π close with #singleton-inv π
... | refl = close
#process π (fail p) with #one+ π p
... | Δ′ , q , π′ = fail q
#process π (wait p P) with #one+ π p
... | Δ′ , q , π′ = wait q (#process π′ P)
#process π (select x p P) with #one+ π p
... | Δ′ , q , π′ = select x q (#process (#next π′) P)
#process π (case p P Q) with #one+ π p
... | Δ′ , q , π′ = case q (#process (#next π′) P) (#process (#next π′) Q)
#process π (fork p q P Q) with #one+ π p
... | Δ′ , p′ , π′ with #split π′ q
... | Δ₁ , Δ₂ , q′ , π₁ , π₂ = fork p′ q′ (#process (#next π₁) P) (#process (#next π₂) Q)
#process π (join p P) with #one+ π p
... | Δ′ , q , π′ = join q (#process (#next (#next π′)) P)
#process π (cut d p P Q) with #split π p
... | Δ₁ , Δ₂ , q , π₁ , π₂ = cut d q (#process (#next π₁) P) (#process (#next π₂) Q)
#process π (server p un P) with #one+ π p
... | Δ′ , q , π′ = server q (#un π′ un) (#process (#next π′) P)
#process π (client p P) with #one+ π p
... | Δ′ , q , π′ = client q (#process (#next π′) P)
#process π (weaken p P) with #one+ π p
... | Δ′ , q , π′ = weaken q (#process π′ P)
#process π (contract p P) with #one+ π p
... | Δ′ , q , π′ = contract q (#process (#next (#next π′)) P)
\end{code}

>>>>

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Bool using (Bool; true; false; if_then_else_)
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong; cong₂)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)

@@

We adopt an \emph{intrinsically typed} representation of processes, so that a
process is also a proof derivation showing that the process is well typed. This
choice increases the effort in the definition of the datatypes for processes and
their operational semantics, but pays off in the rest of the formalization for a
number of reasons:

@@

\begin{code}
data Process : Context → Set where
  link      : ∀{A Γ} → Γ ≃ [ A ] + [ dual A ] → Process Γ
  fail      : ∀{Γ Δ} → Γ ≃ ⊤ , Δ → Process Γ
  wait      : ∀{Γ Δ} → Γ ≃ ⊥ , Δ → Process Δ → Process Γ
  close     : Process [ 𝟙 ]
  case      : ∀{A B Γ Δ} → Γ ≃ A & B , Δ →
              Process (A ∷ Δ) → Process (B ∷ Δ) → Process Γ
  select    : ∀{A B Γ Δ} (x : Bool) → Γ ≃ A ⊕ B , Δ →
              Process ((if x then A else B) ∷ Δ) → Process Γ
  join      : ∀{A B Γ Δ} → Γ ≃ A ⅋ B , Δ → Process (B ∷ A ∷ Δ) → Process Γ
  fork      : ∀{A B Γ Δ Γ₁ Γ₂} → Γ ≃ A ⊗ B , Δ → Δ ≃ Γ₁ + Γ₂ →
              Process (A ∷ Γ₁) → Process (B ∷ Γ₂) → Process Γ
  all       : ∀{A Γ Δ} → Γ ≃ `∀ A , Δ →
              ((B : Type) → Process (subst [ B /_] A ∷ Δ)) → Process Γ
  ex        : ∀{A B Γ Δ} → Γ ≃ `∃ A , Δ → Process (subst [ B /_] A ∷ Δ) → Process Γ
  server    : ∀{A Γ Δ} → Γ ≃ `! A , Δ → Un Δ → Process (A ∷ Δ) → Process Γ
  client    : ∀{A Γ Δ} → Γ ≃ `? A , Δ → Process (A ∷ Δ) → Process Γ
  weaken    : ∀{A Γ Δ} → Γ ≃ `? A , Δ → Process Δ → Process Γ
  contract  : ∀{A Γ Δ} → Γ ≃ `? A , Δ → Process (`? A ∷ `? A ∷ Δ) → Process Γ
  cut       : ∀{A Γ Γ₁ Γ₂} → Γ ≃ Γ₁ + Γ₂ →
              Process (A ∷ Γ₁) → Process (dual A ∷ Γ₂) → Process Γ
\end{code}

@@

% \begin{code}
%    fail      : ∀{Γ Δ} (p : Γ ≃ ⊤ , Δ) → Process Γ
% \end{code}

@@

% \begin{code}
%    close     : Process  [ 𝟙 ]
% \end{code}

@@

% \begin{code}
%    wait      : ∀{Γ Δ} (p : Γ ≃ ⊥ , Δ) → Process Δ → Process Γ
% \end{code}

@@

% \begin{code}
%    select    : ∀{A B Γ Δ} (x : Bool) (p : Γ ≃ A ⊕ B , Δ) →
%                Process ((if x then A else B) ∷ Δ) → Process Γ
% \end{code}

We represent the tags $\InTag_1$ and $\InTag_2$ with Agda's boolean values
\AgdaInductiveConstructor{true} and \AgdaInductiveConstructor{false}. In this
way we can use a single constructor \AgdaInductiveConstructor{select} for the
selections $\Select\x{\InTag_i}.P$. According to the semantics of \Calculus, the
channel on which the selection is performed is consumed and the continuation
process uses a fresh continuation channel. As a consequence, the type of the
continuation channel, which is either $A$ or $B$ depending on the value of the
tag $x$, is \emph{prepended} in front of the typing context used to type the
continuation process.

% \begin{code}
%    case      : ∀{A B Γ Δ} (p : Γ ≃ A & B , Δ) →
%                Process (A ∷ Δ) → Process (B ∷ Δ) → Process Γ
% \end{code}

@@

% \begin{code}
%    fork      : ∀{A B Γ Δ Γ₁ Γ₂} (p : Γ ≃ A ⊗ B , Δ) (q : Δ ≃ Γ₁ + Γ₂) →
%                Process (A ∷ Γ₁) → Process (B ∷ Γ₂) → Process Γ
% \end{code}

@@

% \begin{code}
%    join      : ∀{A B Γ Δ} (p : Γ ≃ A ⅋ B , Δ) → Process (B ∷ A ∷ Δ) →
%                Process Γ
% \end{code}

@@

% \begin{code}
%    server    : ∀{A Γ Δ} (p : Γ ≃ `! A , Δ) (un : Un Δ) →
%                Process (A ∷ Δ) → Process Γ
% \end{code}

@@

% \begin{code}
%    client    : ∀{A Γ Δ} (p : Γ ≃ `? A , Δ) → Process (A ∷ Δ) → Process Γ
%    weaken    : ∀{A Γ Δ} (p : Γ ≃ `? A , Δ) → Process Δ → Process Γ
%    contract  : ∀{A Γ Δ} (p : Γ ≃ `? A , Δ) → Process (`? A ∷ `? A ∷ Δ) →
%                Process Γ
% \end{code}

@@

% \begin{code}
%    cut       : ∀{A Γ Γ₁ Γ₂} (p : Γ ≃ Γ₁ + Γ₂) →
%                Process (A ∷ Γ₁) → Process (dual A ∷ Γ₂) → Process Γ

%    all       : ∀{A Γ Δ} (p : Γ ≃ `∀ A , Δ) →
%                ((B : Type) → Process (subst [ B /_] A ∷ Δ)) → Process Γ
%    ex        : ∀{A B Γ Δ} (p : Γ ≃ `∃ A , Δ) →
%                Process (subst [ B /_] A ∷ Δ) → Process Γ
% \end{code}

@@

Permutations preserve process typeability. In a sense we can read this property
as the fact that we can rename channels at will.

\begin{code}
#process : ∀{Γ Δ} → Γ # Δ → Process Γ → Process Δ
\end{code}
\begin{code}[hide]
#process π (link p) with #one+ π p
... | Δ′ , q , π′ with #singleton-inv π′
... | refl = link q
#process π (fail p) with #one+ π p
... | Δ′ , q , π′ = fail q
#process π (wait p P) with #one+ π p
... | Δ′ , q , π′ = wait q (#process π′ P)
#process π close with #singleton-inv π
... | refl = close
#process π (case p P Q) with #one+ π p
... | Δ′ , q , π′ = case q (#process (#next π′) P) (#process (#next π′) Q)
#process π (select x p P) with #one+ π p
... | Δ′ , q , π′ = select x q (#process (#next π′) P)
#process π (join p P) with #one+ π p
... | Δ′ , q , π′ = join q (#process (#next (#next π′)) P)
#process π (fork p q P Q) with #one+ π p
... | Δ′ , p′ , π′ with #split π′ q
... | Δ₁ , Δ₂ , q′ , π₁ , π₂ = fork p′ q′ (#process (#next π₁) P) (#process (#next π₂) Q)
#process π (all p P) with #one+ π p
... | Δ' , q , π' = all q λ B → #process (#next π') (P B)
#process π (ex p P) with #one+ π p
... | Δ' , q , π' = ex q (#process (#next π') P)
#process π (server p un P) with #one+ π p
... | Δ′ , q , π′ = server q (#un π′ un) (#process (#next π′) P)
#process π (client p P) with #one+ π p
... | Δ′ , q , π′ = client q (#process (#next π′) P)
#process π (weaken p P) with #one+ π p
... | Δ′ , q , π′ = weaken q (#process π′ P)
#process π (contract p P) with #one+ π p
... | Δ′ , q , π′ = contract q (#process (#next (#next π′)) P)
#process π (cut p P Q) with #split π p
... | Δ₁ , Δ₂ , q , π₁ , π₂ = cut q (#process (#next π₁) P) (#process (#next π₂) Q)
\end{code}
